\Large {\bf Protocol-Guided Analysis of Post-silicon Traces Under Limited Observability}

### 2015-10-01 paper.tex
Final version fro ISQED
Added a results table in section VI.

Fixed missing references issues.

Given an obsered trace $\rho$ and the set $\vec{F}$ of LPNs,
Algorithm~\ref{algo:compliance} provides a basic procedure for
computing a set of compliant flow execution scenarios.
\begin{algorithm}[ht]
\DontPrintSemicolon
Create an empty scenario $scen$\;
$\mathit{Scen} = \{scen\}$\;
\ForEach {$h, \; 1 \leq h \leq n $} {
	$\mathit{found} \gets {\tt true}$ \;
	$Scen' = \emptyset$\;
	\ForEach {$scen \in Scen$} {
  		\ForEach {$(F_{i,j}, s_{i,j}) \in \mathit{scen_1}$} {
    		\If {$\mathit{accept}(F_{i,j}, s_{i,j}, e_h) = (F_{i,j}, s'_{i,j})$} {
				Let $scen'$ be a copy of $scen$\;
      		$\mathit{scen'} \gets (\mathit{scen'} - (F_{i,j}, s_{i,j})) \cup (F_{i,j}, s'_{i,j})$\;
      		$\mathit{Scen'} \gets \mathit{scen'} \cup \mathit{Scen'}$\;
      		$\mathit{found} \gets {\tt false}$ \;
    		}
  		}
  		\ForEach {$F_i \in \vec{F}$} {
      	create a new instance $F_{i, j+1}$ \;
      	\If {$\mathit{accept}(F_{i,j+1},\mathit{init}_{i,j+1}, e_h) = (F_{i,j+1}, s'_{i,j+1})$} {
    			Let $scen'$ be a copy of $scen$\;
				$\mathit{scen'} \gets \mathit{scen'} \cup (F_{i,j+1}, s'_{i,j+1})$ \;
				$\mathit{Scen'} \gets \mathit{scen'} \cup \mathit{Scen'}$\;
        		$\mathit{found} \gets {\tt false}$ \;
      	}
    	}
	}
  \If {$\mathit{found} == {\tt true}$} {
    \Return ${\tt Inconsistent}$\;
  }
  $Scen = Scen'$\;
}
\Return $\mathit{scen}$ \;
\caption{$\textsc{Check-Compliance}(\vec{F}, \, \rho)$}
\label{algo:compliance}
\end{algorithm}
The algorithm operates by keeping track (in variable
$\mathit{Scen}$) of a set of candidate flow execution scenarios
compliant with each prefix of $\rho$.  At each iteration,
for each event $e_h$ in the observed trace, we
update $Scen$ by either extending a member of
$\mathit{scen}$ or initiating a new protocol instance for 
each $scen \in Scen$ with respect to $e_h$ in every possible way.
If $e_h$ cannot be emitted by any existing or new flow instances, 
then we report that the trace is {\tt  inconsistent}, \ie, 
there is no possible interleaving of
the protocol instances from $\vec{F}$ that is compliant with
$\rho$.

@@

After the SoC model finishes executing the program, there
are totally $343581$ messages collected in the trace file.
Not all of the messages are relevant to the flow
specification as many are used by GEM5 to initialize its
simulation environment.  After removing those irrelevant
messages, the number of messages in the trace file is to
reduced to $121138$.

@@

To take the partial observability into account, the four
monitors attached to the links between two CPUs and their
caches are disabled. Then, the trace is generated by the
remaining five monitors from the SoC model executing the
same program. After removing all non-observable messages,
the trace only contains 15089 messages. The numbers of the
flow instances extracted by the trace analysis method are
shown in Table~\ref{table-par-obs}.  From these results, the
numbers of the flow instances are dropped significantly
compared to the results extracted from the trace with the
full observability as shown in
Table~\ref{table-case-2}. This difference is due to that
some communications occurred in the system when executing
the program involve the CPUs and their corresponding caches
only, and the traffic on the links between the CPUs and
their corresponding caches is not observable. Therefore, the
instances of the flow specifications characterizing these
communications do not exist in the trace. In other words,
all extracted flow instances in Table~\ref{table-par-obs}
characterize the communications that pass through the memory
bus in the system model.  The runtime and memory usage is
similar to that for analyzing the trace of the full
observability as shown above.

@@

The whole experiment takes about $15$ minutes and $420$~MB to finish, significantly higher than the numbers for analyzing traces where there is no ambiguity in the observed events.  This is due to the fact that a trace of ambiguous events is in fact a set of traces of original events, which lead to large numbers of execution scenarios either during or at the end of the analysis.  In this experiment, the peak number of execution scenarios during the analysis process is $70384$.   Compared to the final number, many of the intermediate execution scenarios are invalid, and removed eventually.  However, controlling the number of intermediate execution scenarios during the trace analysis is critical in order for the analysis to be tractable.  Here, insights from validators can help.

As shown above, the ambiguous interpretation of events can lead to large numbers of intermediate and final execution scenarios, which not only make the trace analysis more time consuming but also make it difficult to gain insightful understanding from the derived execution scenarios.  Careful selection of what to observe may have big impact on results from the trace analysis.  In this last experiment, we relax the assumption made in the previous experiment such that the events passing each link are partitioned into two groups, one for read operations and one for write operations.  Similar to the assumption made in the previous experiment, events in the same group are assumed to be non-distinguishable.  The monitors are modified accordingly such that they output all events in the same group into the trace file if an event from that group is captured.  After the trace analysis on this new partially observed trace is finished, only one execution scenario is derived where the distribution of the numbers of flow instances is the same as those shown in Table~\ref{table-par-obs}.  The peak number of execution scenarios encountered during the trace analysis is $4$.  The total runtime is about $1$~second.  Compared to the results from the previous experiment, the precision and the performance of the trace analysis are improved dramatically as a result of careful selection of observable signals.

>>>>

Given an obsered trace $\rho$ and the set $\vec{F}$ of LPNs,
Algorithm~\ref{algo:compliance} provides a basic procedure for
computing a set of compliant flow execution scenarios.
\begin{algorithm}[ht]
\DontPrintSemicolon
Create an empty scenario $scen$\;
$\mathit{Scen} = \{scen\}$\;
\ForEach {$h, \; 1 \leq h \leq n $} {
	$\mathit{found} \gets {\tt true}$ \;
	$Scen' = \emptyset$\;
	\ForEach {$scen \in Scen$} {
  		\ForEach {$(F_{i,j}, s_{i,j}) \in \mathit{scen_1}$} {
    		\If {$\mathit{accept}(F_{i,j}, s_{i,j}, e_h) = (F_{i,j}, s'_{i,j})$} {
				Let $scen'$ be a copy of $scen$\;
      		$\mathit{scen'} \gets (\mathit{scen'} - (F_{i,j}, s_{i,j})) \cup (F_{i,j}, s'_{i,j})$\;
      		$\mathit{Scen'} \gets \mathit{scen'} \cup \mathit{Scen'}$\;
      		$\mathit{found} \gets {\tt false}$ \;
    		}
  		}
  		\ForEach {$F_i \in \vec{F}$} {
      	create a new instance $F_{i, j+1}$ \;
      	\If {$\mathit{accept}(F_{i,j+1},\mathit{init}_{i,j+1}, e_h) = (F_{i,j+1}, s'_{i,j+1})$} {
    			Let $scen'$ be a copy of $scen$\;
				$\mathit{scen'} \gets \mathit{scen'} \cup (F_{i,j+1}, s'_{i,j+1})$ \;
				$\mathit{Scen'} \gets \mathit{scen'} \cup \mathit{Scen'}$\;
        		$\mathit{found} \gets {\tt false}$ \;
      	}
    	}
	}
  \If {$\mathit{found} == {\tt true}$} {
    \Return ${\tt Inconsistent}$\;
  }
  $Scen = Scen'$\;
}
\Return $\mathit{Scen}$ \;
\caption{$\textsc{Check-Compliance}(\vec{F}, \, \rho)$}
\label{algo:compliance}
\end{algorithm}
The algorithm operates by keeping track (in variable
$\mathit{Scen}$) of a set of candidate flow execution scenarios
compliant with each prefix of $\rho$.  At each iteration,
for each event $e_h$ in the observed trace, we
update $Scen$ by either extending a member of
$\mathit{scen}$ or initiating a new protocol instance for 
each $scen \in Scen$ with respect to $e_h$ in every possible way.
If $e_h$ cannot be emitted by any existing or new flow instances, 
then we report that the trace is {\tt  inconsistent}, \ie, 
there is no possible interleaving of
the protocol instances from $\vec{F}$ that is compliant with
$\rho$.

@@

\begin{table}[tb]
\caption{Runtime Results of Trace analysis. Time is in seconds and memory usage is in MB.}
\begin{center}
\begin{tabular}{|c|c|c|c|c|}
\hline
	&	F-Obs. & \begin{tabular}{c}P-Obs. \\ No Amb.\end{tabular} & \begin{tabular}{c}P-Obs. \\ Amb.~1\end{tabular} & \begin{tabular}{c}P-Obs. \\ Amb.~2\end{tabular} \\
\hline
\hline
Time & $3$ & $2.78$ & $896$ & $<1$\\
\hline
Mem & $12$ & $10$ & $420$ & $9$\\
\hline

\end{tabular}
\end{center}
\label{table-results}
\end{table}%

After this model is executed with the simple concurrent programs, 
the trace analysis is applied to traces with different observabilities
collected from this model.  The runtime results are shown in Table~\ref{table-results}.
The first column shows the results from analyzing the trace with the full
observability, while the next three show the result from 
analyzing traces with different partial observability assumptions.

In the first experiment, full observability is assumed.
After the SoC model finishes executing the program, there
are totally $343581$ messages collected in the trace file.
Not all of the messages are relevant to the flow
specification as many are used by GEM5 to initialize its
simulation environment.  After removing those irrelevant
messages, the number of messages in the trace file is to
reduced to $121138$.

@@

In the second experiment, partial observability is taken into account
with the four monitors attached to the links between two CPUs and their
caches are disabled. Then, the trace is generated by the
remaining five monitors from the SoC model executing the
same program. After removing all non-observable messages,
the trace only contains 15089 messages. The numbers of the
flow instances extracted by the trace analysis method are
shown in Table~\ref{table-par-obs}.  From these results, the
numbers of the flow instances are dropped significantly
compared to the results extracted from the trace with the
full observability as shown in
Table~\ref{table-case-2}. This difference is due to that
some communications occurred in the system when executing
the program involve the CPUs and their corresponding caches
only, and the traffic on the links between the CPUs and
their corresponding caches is not observable. Therefore, the
instances of the flow specifications characterizing these
communications do not exist in the trace. In other words,
all extracted flow instances in Table~\ref{table-par-obs}
characterize the communications that pass through the memory
bus in the system model.  The runtime and memory usage as shown in
the third column in Table~\ref{table-results} are
similar to those for analyzing the trace of the full
observability.

@@

The whole experiment takes about $15$ minutes and $420$~MB to finish as shown in column~4
in Table~\ref{table-results}, significantly higher than the numbers for analyzing traces where there is no ambiguity in the observed events.  This is due to the fact that a trace of ambiguous events is in fact a set of traces of original events, which lead to large numbers of execution scenarios either during or at the end of the analysis.  In this experiment, the peak number of execution scenarios during the analysis process is $70384$.   Compared to the final number, many of the intermediate execution scenarios are invalid, and removed eventually.  However, controlling the number of intermediate execution scenarios during the trace analysis is critical in order for the analysis to be tractable.  Here, insights from validators can help.

As shown above, the ambiguous interpretation of events can lead to large numbers of intermediate and final execution scenarios, which not only make the trace analysis more time consuming but also make it difficult to gain insightful understanding from the derived execution scenarios.  Careful selection of what to observe may have big impact on results from the trace analysis.  In this last experiment, we relax the assumption made in the previous experiment such that the events passing each link are partitioned into two groups, one for read operations and one for write operations.  Similar to the assumption made in the previous experiment, events in the same group are assumed to be non-distinguishable.  The monitors are modified accordingly such that they output all events in the same group into the trace file if an event from that group is captured.  After the trace analysis on this new partially observed trace is finished, only one execution scenario is derived where the distribution of the numbers of flow instances is the same as those shown in Table~\ref{table-par-obs}.  The peak number of execution scenarios encountered during the trace analysis is $4$.  The total runtime and memory usage are negligible as shown in the last column in Table~\ref{table-results}.  Compared to the results from the previous experiment, the precision and the performance of the trace analysis are improved dramatically as a result of careful selection of observable events.
